\paperTitle

### 2014-05-12 doc/icml2014/directed.tex
Updated sigma_max, sigma_min to be sigma_1, sigma_k
Updated uniformly bottlenecked and the non-degeneracy assumption

%Let us now generalize the intuitions of the directed grid to general graphs.
The intuitions of the directed grid generalize readily to general graphs.
% PL: this prose is redundant with property
%Note that in each of the above cases,
  %for any clique $\sC$, each hidden variable $h_i \in \sC$ has its own
  %view which is conditionally independent given the other hidden variables in
  %the clique.
%Fundamentally, algorithm takes advantage of having a view for every
  %hidden variable in a clique to help identify its hidden state.
The key property required for our algorithm to work is as follows:
%We capture this general property as follows:
\begin{property}[Exclusive views]
  \label{prop:exclusive-views}
A hidden clique $\sC \in \sG$ is said to possess the \emph{exclusive views property} if for
  every hidden variable $h_i \in \sC$, (i)
  there exists some observed variable,
  $x_{v}$ which is conditionally independent of the rest of the clique
  given $h_i$, and (ii) the conditional moments $\mOpp{v}{i} \eqdef
  \Pr(x_{v} \mid h_i)$ can be estimated consistently.
We call $x_v$ an exclusive view of $h_i$ in $\sC$.
If every clique has the exclusive views property, then we say that $\sG$ does as well.
\end{property}
% PL: to save space
%This property also extends to any connected subgraph $\sG' \in \sG$.
%For the rest of this section, we will only consider subgraphs which are
  %cliques. 
%In \sectionref{undirected}, we will show how the generalization
  %to subgraphs allows us to efficiently estimate the pseudo-likelihood.

@@

% DONE: don't talk about parameter recovery yet, just establish exclusive views
%Finally, the claim we make is that we can recover parameters for any
  %graphical model where every hidden variable is a bottleneck:
Now define the following property:
\begin{property}[Uniformly bottlenecked]
  \label{prop:bottleneck}
  A graphical model $\sG$ is \emph{uniformly bottlenecked} if
  every hidden variable in $\sG$ is a bottleneck.
\end{property}
Note that two hidden variables are allowed to have overlapping views.
%If every hidden variable is a bottleneck, then we can estimate $\mOpp{v}{i}$
It is clear that having uniform bottlenecks implies part (ii)
of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
but we will show that part (i) follows as well.

@@

To apply $\TensorFactorize$, we need that each $\mOpp{v}{i}$ has full
  column rank and that each $\pi^{(i)} \eqdef \Pr(h_i) \succ 0$.
% But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to
%   $x_v$, marginalizing out the rest of the graph.
% To make this property more interpretable, let us assume the following
%   property which we prove to be sufficient in
%   \appendixref{assumption-proof}:
% \begin{assumption} 
% \label{asm:full-rank-plus}
% The marginal distribution of every latent variable has full support:
%   $\forall{h_i \in H},~ \mPi{i} \succ 0$.
% For every clique $\sC \in \sG$ (including ones with observed variables),
%   and every conditional distribution $\Pr(h_c \given \Pa(h_c))$, where
%   $h_c \in \sC$, every {\em tensor unfolding} has full column rank, $k$.
% \end{assumption}
% In our directed grid example (\sectionref{directedExample}), for the
%   clique $\sC = \{ h_1, h_2 \}$ this condition implies $\pi = \Pr(h_1)
%   \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; for the clique
%   $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column
%   rank $k$.

@@

From \citet{anandkumar13tensor} given \assumptionref{full-rank}, we
  have that $\mOpphat{v}{i}$ converges to $\mOpp{v}{i}$ at a rate of
  $n^{-\frac12}$ with a constant that depends polynomially on the $k$-th
  singular value of $\mOpp{v}{i}$:
\begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
  \label{thm:sample-complexity}
  Wlog, let $v=1$ and $i=1$. Let $x_1, x_2, x_3$ be three views for
  $h_1$. With probability at least $1 - \delta$,
\begin{align*}
  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
    &\le  \\
    &
      O\left(
      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min} } 
      {\sqrt{n}(\sigma_{\min}(\mOpp{1}{1}) \sigma_{\min}(\mOpp{1}{2}))^{5/2}} \right) \epsilon.
\end{align*}

@@

In general, $\sigma_{\min}(\mOpp{v}{i})$ can depend exponentially on the
  path length between the hidden variable $h_i$ and $x_v$.
This is expected because the influence from $h_i$ to $x_v$
  gets diluted for every latent variable it must factor through.

>>>>

%Let us now generalize the intuitions of the directed grid to general graphs.
The intuitions of the directed grid generalize readily to general graphs.
% PL: this prose is redundant with property
%Note that in each of the above cases,
  %for any clique $\sC$, each hidden variable $h_i \in \sC$ has its own
  %view which is conditionally independent given the other hidden variables in
  %the clique.
%Fundamentally, algorithm takes advantage of having a view for every
  %hidden variable in a clique to help identify its hidden state.
The key property required for our algorithm to work is as follows:
%We capture this general property as follows:
\begin{property}[Exclusive views]
  \label{prop:exclusive-views}
A hidden clique $\sC \in \sG$ is said to possess the \emph{exclusive views property} if for
  every hidden variable $h_i \in \sC$, (i)
  there exists some observed variable,
  $x_{v}$ which is conditionally independent of the rest of the clique
  given $h_i$, and (ii) the conditional moments $\mOpp{v}{i} \eqdef
  \Pr(x_{v} \mid h_i)$ can be estimated consistently.
We call $x_v$ an exclusive view of $h_i$ in $\sC$.
If every clique has the exclusive views property, then we say that $\sG$ does as well.
\end{property}

Next, we state the non-degeneracy assumptions under which our general
  algorithm is guaranteed to obtain consistent estimates.
\begin{assumption}[Non-degeneracy]
  \label{asm:non-degeneracy}
  (i) The marginal distribution of each hidden variable $h_i$ have
  full support: $\mPi{i} \succ 0$. 
  (ii) The conditional moments $\mOpp{v}{i}$ have full column rank $k$.
\end{assumption}

% PL: to save space
%This property also extends to any connected subgraph $\sG' \in \sG$.
%For the rest of this section, we will only consider subgraphs which are
  %cliques. 
%In \sectionref{undirected}, we will show how the generalization
  %to subgraphs allows us to efficiently estimate the pseudo-likelihood.

@@

% DONE: don't talk about parameter recovery yet, just establish exclusive views
%Finally, the claim we make is that we can recover parameters for any
  %graphical model where every hidden variable is a bottleneck:
Now define the following property:
\begin{property}[Uniformly bottlenecked]
  \label{prop:bottleneck}
  A clique $\sC$ is \emph{uniformly bottlenecked} if
  every hidden variable in $\sC$ is a bottleneck. 
  If every clique $\sC \in \sG$ is uniformly bottlenecked, then we say
  that the graph $\sG$ is uniformly bottlenecked as well.
\end{property}
Note that two hidden variables are allowed to have overlapping views.
%If every hidden variable is a bottleneck, then we can estimate $\mOpp{v}{i}$
It is clear that having uniform bottlenecks implies part (ii)
of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
but we will show that part (i) follows as well.

@@

% But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to
%   $x_v$, marginalizing out the rest of the graph.
% To make this property more interpretable, let us assume the following
%   property which we prove to be sufficient in
%   \appendixref{assumption-proof}:
% \begin{assumption} 
% \label{asm:full-rank-plus}
% The marginal distribution of every latent variable has full support:
%   $\forall{h_i \in H},~ \mPi{i} \succ 0$.
% For every clique $\sC \in \sG$ (including ones with observed variables),
%   and every conditional distribution $\Pr(h_c \given \Pa(h_c))$, where
%   $h_c \in \sC$, every {\em tensor unfolding} has full column rank, $k$.
% \end{assumption}
% In our directed grid example (\sectionref{directedExample}), for the
%   clique $\sC = \{ h_1, h_2 \}$ this condition implies $\pi = \Pr(h_1)
%   \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; for the clique
%   $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column
%   rank $k$.

@@

From \citet{anandkumar13tensor} given the non-degeneracy assumptions
  \assumptionref{non-degeneracy}, we have that $\mOpphat{v}{i}$
  converges to $\mOpp{v}{i}$ at a rate of $n^{-\frac12}$ with a constant
  that depends polynomially on the $k$-th singular value of
  $\mOpp{v}{i}$:
\begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
  \label{thm:sample-complexity}
  Wlog, let $v=1$ and $i=1$. Let $x_1, x_2, x_3$ be three views for
  $h_1$. With probability at least $1 - \delta$,
\begin{align*}
  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
    &\le  \\
    &
      O\left(
      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min} } 
      {\sqrt{n}(\sigma_{k}(\mOpp{1}{1}) \sigma_{k}(\mOpp{1}{2}))^{5/2}} \right) \epsilon.
\end{align*}

@@

In general, $\sigma_{k}(\mOpp{v}{i})$ can depend exponentially on the
  path length between the hidden variable $h_i$ and $x_v$.
This is expected because the influence from $h_i$ to $x_v$
  gets diluted for every latent variable it must factor through.
